The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-01-21 paper/ms.tex
Added table with companion data
Updated paper

\begin{abstract}
Young, intermediate-mass stars are experiencing renewed interest as targets for direct-imaging planet searches. However, these types of stars are part of multiple systems more often than not. Close stellar companions affect the formation and orbital evolution of any planets, and the properties of the companions can help constrain the binary formation mechanism. Unfortunately, close companions are difficult and expensive to detect with imaging techniques. In this paper, we describe the direct spectral detection method wherein a high-resolution spectrum of the primary is cross-correlated against a template for a companion star. Variants of this method have previously been used to search for stellar, brown dwarf, and even planetary companions. We show that the direct spectral detection method can detect companions as late as M-type orbiting A0 or earlier primary stars in a single epoch on small-aperture telescopes. In addition to estimating the detection limits, we determine the sources of uncertainty in characterizing the companion temperature, and find that large systematic biases can exist. After calibrating the systematic biases with synthetic binary star observations, we apply the method to a sample of 34 known binary systems with an A- or B-type primary star. We detect nine total companions, including four of the five known companions with literature temperatures between $4000$ K $ < T < 6000$ K, the temperature range for which our method is optimized. We additionally characterize the companion for the first time in two previously single-lined binary systems and one binary identified with speckle interferometry. This method provides an inexpensive way to use small-aperture telescopes to detect binary companions with moderate mass-ratios, and is competitive with high-resolution imaging techniques inside $\sim 100-200$ mas.
\end{abstract}

@@

\label{sec:intro}
There has recently been a revival of interest in the multiplicity properties of intermediate-mass stars,  spurred largely by the detection of planets orbiting nearby \mbox{$\sim2 M_{\odot}$} stars on both wide \citep[e.g.][]{Lagrange2010, Marois2008} and close \citep{Johnson2011} orbits. In this context, binary companions are contaminants; companions complicate radial velocity planet searches because they necessitate simultaneous modeling of both stellar motions \citep[e.g.][]{Bergmann2015}. Likewise, companions complicate direct imaging planet searches by requiring either extremely high-contrast instrumentation \citep{Thalmann2014} or specialized coronagraphs \citep{Crepp2010}.

However, known binary stars are typically avoided in planet search programs for a more fundamental reason: the binary companion depletes or destroys the planet-forming disk. By combining a binary census of the $\sim 2$ Myr Taurus-Auriga star-forming region with a disk census of the same, \cite{Kraus2012} showed that close ($\lesssim 40$ AU) binaries are about 2-3 times less likely to host a protoplanetary disk, and so hasten disk dispersal. Even if a disk survives, it tends to be depleted in mass by a factor of $\sim 25$ for binary separations $\lesssim 30$ AU \citep{Harris2012}.
%[CITE KRAUS PAPER] performed a search for binary companions among Kepler planet host stars, and found that the presence of a companion within 5-50AU suppresses planet formation by a factor of $\sim 4$.

@@

>>>>

\begin{abstract}
Abstract here
\end{abstract}

@@

\label{sec:intro}
%There has recently been a revival of interest in the multiplicity properties of intermediate-mass stars,  spurred largely by the detection of planets orbiting nearby \mbox{$\sim2 M_{\odot}$} stars on both wide \citep[e.g.][]{Lagrange2010, Marois2008} and close \citep{Johnson2011} orbits. In this context, binary companions are contaminants; companions complicate radial velocity planet searches because they necessitate simultaneous modeling of both stellar motions \citep[e.g.][]{Bergmann2015}. Likewise, companions complicate direct imaging planet searches by requiring either extremely high-contrast instrumentation \citep{Thalmann2014} or specialized coronagraphs \citep{Crepp2010}.

%However, known binary stars are typically avoided in planet search programs for a more fundamental reason: the binary companion depletes or destroys the planet-forming disk. By combining a binary census of the $\sim 2$ Myr Taurus-Auriga star-forming region with a disk census of the same, \cite{Kraus2012} showed that close ($\lesssim 40$ AU) binaries are about 2-3 times less likely to host a protoplanetary disk, and so hasten disk dispersal. Even if a disk survives, it tends to be depleted in mass by a factor of $\sim 25$ for binary separations $\lesssim 30$ AU \citep{Harris2012}.
%[CITE KRAUS PAPER] performed a search for binary companions among Kepler planet host stars, and found that the presence of a companion within 5-50AU suppresses planet formation by a factor of $\sim 4$.

Stellar multiplicity is an inevitable and common outcome of star formation, with roughly half of all solar-type field stars in binary or multiple systems \citep{Raghavan2010} and an even higher fraction as the stellar mass increases. Young stellar associations and clusters tend to have even higher multiplicity \citep{Duchene2013}, indicating that stars often form in multiple systems that are destroyed by dynamical interactions as the cluster dissociates.

TODO: mass-ratio is the most dynamically stable orbital feature. Where did I see this?

TODO: \citep{Carrier2002} and \citep{Vuissoz2004} look at Am/Ap stars and find a mass-ratio distribution with a broad peak near $q \sim 0.5$. This is coming mostly from SB1s, and assuming uniform inclinations to go from the mass function to the companion mass, and make simple assumptions about the primary mass. Still, it is interesting that my result is pretty similar for normal A-stars.

@@

\begin{longtable*}{|l|lrrll|}
\hline
       star &      $T_\mathrm{eff}$ &   $v\sin{i}$ &  [Fe/H] &   Isochrone Mass &   Spectral Type Mass \\
\hline
\endhead
\hline
\multicolumn{3}{r}{{Continued on next page}} \\
\endfoot
   HIP 2548 &  $5732 \pm 112$ &      30 &    -0.5 &  $1.0^{+0.06}_{-0.04}$ &  $1.0^{+0.04}_{-0.03}$ \\
   HIP 3478 &  $5524 \pm 150$ &       5 &    -0.5 &  $1.1^{+0.28}_{-0.13}$ &  $1.0^{+0.03}_{-0.04}$ \\
   HIP 5310 &  $3343 \pm 154$ &      30 &     0.0 &  $0.3^{+0.13}_{-0.08}$ &  $0.4^{+0.12}_{-0.12}$ \\
     HR 545 &   $4312 \pm 87$ &       5 &    -0.5 &  $0.8^{+0.11}_{-0.10}$ &  $0.7^{+0.01}_{-0.02}$ \\
     HR 604 &  $4736 \pm 150$ &       5 &     0.5 &    $nan^{+nan}_{-nan}$ &  $0.8^{+0.02}_{-0.02}$ \\
  HIP 10732 &  $5578 \pm 109$ &       5 &     0.0 &  $1.0^{+0.29}_{-0.04}$ &  $1.0^{+0.02}_{-0.04}$ \\
  HIP 12332 &  $5551 \pm 107$ &      10 &     0.0 &    $nan^{+nan}_{-nan}$ &  $1.0^{+0.01}_{-0.04}$ \\
  HIP 12706 &  $6051 \pm 107$ &      10 &     0.0 &  $1.1^{+0.03}_{-0.03}$ &  $1.1^{+0.04}_{-0.05}$ \\
  HIP 13165 &  $5773 \pm 163$ &       5 &    -0.5 &  $1.1^{+0.31}_{-0.08}$ &  $1.0^{+0.10}_{-0.03}$ \\
  HIP 13327 &  $5524 \pm 150$ &       5 &    -0.5 &  $1.2^{+0.20}_{-0.21}$ &  $1.0^{+0.03}_{-0.04}$ \\
  HIP 14576 &  $6980 \pm 103$ &      10 &     0.0 &  $1.4^{+0.03}_{-0.03}$ &  $1.5^{+0.05}_{-0.03}$ \\
  HIP 14764 &  $5601 \pm 108$ &      30 &     0.5 &    $nan^{+nan}_{-nan}$ &  $1.0^{+0.03}_{-0.03}$ \\
  HIP 16244 &  $5592 \pm 107$ &       5 &    -0.5 &  $1.2^{+0.20}_{-0.22}$ &  $1.0^{+0.03}_{-0.03}$ \\
  HIP 16340 &  $6214 \pm 150$ &       5 &    -0.5 &  $1.2^{+0.29}_{-0.07}$ &  $1.2^{+0.06}_{-0.05}$ \\
  HIP 16611 &  $5476 \pm 154$ &       1 &     0.5 &  $0.9^{+0.04}_{-0.04}$ &  $0.9^{+0.04}_{-0.03}$ \\
  HIP 17563 &  $8574 \pm 147$ &       5 &     0.5 &  $1.9^{+0.25}_{-0.11}$ &  $2.0^{+0.05}_{-0.06}$ \\
  HIP 18396 &  $4492 \pm 162$ &      30 &    -0.5 &  $0.9^{+0.10}_{-0.12}$ &  $0.8^{+0.03}_{-0.02}$ \\
  HIP 18724 &  $6831 \pm 154$ &      20 &     0.0 &  $1.7^{+0.13}_{-0.23}$ &  $1.4^{+0.04}_{-0.04}$ \\
  HIP 19949 &  $6492 \pm 155$ &      20 &    -0.5 &  $1.3^{+0.06}_{-0.05}$ &  $1.3^{+0.06}_{-0.06}$ \\
  HIP 20380 &  $6492 \pm 109$ &      20 &    -0.5 &  $1.3^{+0.04}_{-0.04}$ &  $1.3^{+0.03}_{-0.05}$ \\
  HIP 21589 &  $4562 \pm 154$ &       5 &     0.0 &  $0.7^{+0.04}_{-0.03}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 22833 &  $4745 \pm 107$ &      10 &     0.0 &  $0.8^{+0.02}_{-0.02}$ &  $0.8^{+0.02}_{-0.01}$ \\
  HIP 22958 &  $6126 \pm 112$ &      30 &    -0.5 &  $1.4^{+0.22}_{-0.21}$ &  $1.2^{+0.04}_{-0.04}$ \\
  HIP 23362 &  $6283 \pm 158$ &       5 &    -0.5 &  $1.2^{+0.06}_{-0.05}$ &  $1.2^{+0.07}_{-0.05}$ \\
  HIP 23916 &  $5450 \pm 158$ &       5 &    -0.5 &  $1.1^{+0.20}_{-0.20}$ &  $0.9^{+0.04}_{-0.03}$ \\
  HIP 24244 &  $4929 \pm 158$ &       5 &    -0.5 &  $0.8^{+0.04}_{-0.04}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 24902 &  $5679 \pm 154$ &      30 &     0.0 &  $1.0^{+0.05}_{-0.04}$ &  $1.0^{+0.04}_{-0.03}$ \\
  HIP 25143 &  $5883 \pm 109$ &      30 &     0.0 &  $1.1^{+0.04}_{-0.03}$ &  $1.1^{+0.05}_{-0.07}$ \\
  HIP 26126 &   $5841 \pm 91$ &       5 &    -0.5 &  $1.0^{+0.03}_{-0.03}$ &  $1.0^{+0.05}_{-0.04}$ \\
  HIP 32607 &   $5068 \pm 91$ &       5 &    -0.5 &  $0.8^{+0.02}_{-0.02}$ &  $0.9^{+0.01}_{-0.01}$ \\
  HIP 37450 &   $nan \pm nan$ & \nodata & \nodata &    $nan^{+nan}_{-nan}$ &    $nan^{+nan}_{-nan}$ \\
  HIP 38538 &  $6576 \pm 150$ &      30 &    -0.5 &  $1.3^{+0.05}_{-0.05}$ &  $1.3^{+0.10}_{-0.05}$ \\
  HIP 38593 &  $5398 \pm 112$ &       5 &    -0.5 &  $1.3^{+0.08}_{-0.13}$ &  $0.9^{+0.03}_{-0.03}$ \\
  HIP 38846 &  $6595 \pm 158$ &      10 &     0.0 &  $1.7^{+0.12}_{-0.21}$ &  $1.3^{+0.09}_{-0.05}$ \\
  HIP 39847 &  $5750 \pm 107$ &      10 &    -0.5 &  $1.0^{+0.04}_{-0.03}$ &  $1.0^{+0.05}_{-0.03}$ \\
  HIP 40881 &  $4562 \pm 154$ &      10 &    -0.5 &  $0.8^{+0.18}_{-0.05}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 42129 &  $5241 \pm 158$ &      10 &    -0.5 &  $1.2^{+0.12}_{-0.18}$ &  $0.9^{+0.04}_{-0.02}$ \\
  HIP 46283 &   $9268 \pm 85$ &      10 &     0.5 &    $nan^{+nan}_{-nan}$ &  $2.2^{+0.02}_{-0.02}$ \\
  HIP 51362 &  $5033 \pm 158$ &       5 &    -0.5 &  $0.8^{+0.04}_{-0.04}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 52678 &  $5033 \pm 112$ &      10 &    -0.5 &  $1.0^{+0.16}_{-0.19}$ &  $0.8^{+0.01}_{-0.02}$ \\
  HIP 55266 &  $8483 \pm 139$ &      30 &    -0.5 &  $1.8^{+0.06}_{-0.04}$ &  $2.0^{+0.06}_{-0.05}$ \\
  HIP 63945 &  $7038 \pm 162$ &      10 &    -0.5 &  $1.6^{+0.26}_{-0.16}$ &  $1.5^{+0.08}_{-0.06}$ \\
  HIP 72154 &  $4918 \pm 162$ &      30 &    -0.5 &  $0.9^{+0.22}_{-0.09}$ &  $0.8^{+0.03}_{-0.03}$ \\
  HIP 76267 &  $5449 \pm 158$ &       5 &    -0.5 &  $0.9^{+0.04}_{-0.04}$ &  $0.9^{+0.04}_{-0.03}$ \\
  HIP 77336 &  $6407 \pm 112$ &      20 &     0.0 &  $1.2^{+0.04}_{-0.04}$ &  $1.3^{+0.04}_{-0.05}$ \\
  HIP 77516 &  $6825 \pm 141$ &       5 &     0.5 &  $1.4^{+0.21}_{-0.06}$ &  $1.4^{+0.03}_{-0.04}$ \\
  HIP 78820 &  $5658 \pm 158$ &       5 &    -0.5 &  $1.3^{+0.11}_{-0.13}$ &  $1.0^{+0.04}_{-0.04}$ \\
  HIP 78821 &  $4968 \pm 154$ &      10 &    -0.5 &  $1.1^{+0.09}_{-0.10}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 79199 &  $4616 \pm 158$ &       5 &    -0.5 &  $0.9^{+0.14}_{-0.14}$ &  $0.8^{+0.02}_{-0.03}$ \\
  HIP 79404 &  $4773 \pm 112$ &      10 &    -0.5 &  $1.1^{+0.07}_{-0.08}$ &  $0.8^{+0.02}_{-0.01}$ \\
  HIP 80460 &  $6391 \pm 155$ &      10 &     0.0 &  $1.2^{+0.06}_{-0.05}$ &  $1.3^{+0.06}_{-0.06}$ \\
  HIP 82673 &  $4668 \pm 112$ &       5 &     0.5 &  $0.8^{+0.03}_{-0.03}$ &  $0.8^{+0.01}_{-0.02}$ \\
  HIP 84606 &  $5476 \pm 154$ &      10 &    -0.5 &  $0.9^{+0.05}_{-0.04}$ &  $0.9^{+0.04}_{-0.03}$ \\
  HIP 85379 &  $7590 \pm 151$ &      20 &     0.0 &    $nan^{+nan}_{-nan}$ &  $1.7^{+0.05}_{-0.03}$ \\
  HIP 85998 &   $5460 \pm 91$ &       1 &    -0.5 &  $1.0^{+0.28}_{-0.03}$ &  $0.9^{+0.04}_{-0.01}$ \\
  HIP 88116 &  $5293 \pm 112$ &       5 &     0.0 &  $0.9^{+0.03}_{-0.03}$ &  $0.9^{+0.02}_{-0.03}$ \\
  HIP 88818 &   $5296 \pm 91$ &      20 &    -0.5 &  $0.9^{+0.02}_{-0.02}$ &  $0.9^{+0.02}_{-0.03}$ \\
  HIP 88817 &  $3647 \pm 154$ &      20 &     0.0 &  $0.5^{+0.07}_{-0.10}$ &  $0.5^{+0.03}_{-0.05}$ \\
  HIP 91118 &  $6492 \pm 155$ &      10 &    -0.5 &  $1.3^{+0.26}_{-0.07}$ &  $1.3^{+0.06}_{-0.06}$ \\
  HIP 93805 &  $4054 \pm 109$ &       5 &    -0.5 &  $0.6^{+0.03}_{-0.03}$ &  $0.6^{+0.04}_{-0.02}$ \\
  HIP 96840 &  $9942 \pm 152$ &      10 &     0.0 &    $nan^{+nan}_{-nan}$ &  $2.3^{+0.03}_{-0.01}$ \\
 HIP 100221 &  $5273 \pm 154$ &      20 &    -0.5 &  $0.9^{+0.29}_{-0.07}$ &  $0.9^{+0.04}_{-0.03}$ \\
 HIP 100907 &  $5222 \pm 109$ &      10 &     0.0 &  $0.9^{+0.03}_{-0.03}$ &  $0.9^{+0.04}_{-0.02}$ \\
 HIP 103298 &   $4615 \pm 90$ &       5 &    -0.5 &  $0.7^{+0.02}_{-0.02}$ &  $0.8^{+0.01}_{-0.02}$ \\
 HIP 109139 &  $5710 \pm 112$ &      20 &    -0.5 &  $1.0^{+0.04}_{-0.03}$ &  $1.0^{+0.03}_{-0.03}$ \\
 HIP 109521 &   $4412 \pm 87$ &      10 &     0.0 &  $0.7^{+0.02}_{-0.02}$ &  $0.7^{+0.01}_{-0.01}$ \\
 HIP 115115 &   $4943 \pm 91$ &       5 &    -0.5 &  $0.8^{+0.02}_{-0.02}$ &  $0.8^{+0.02}_{-0.02}$ \\
 \label{tab:companions}
